---
title: "Bismarck Replication Project"
output: github_document
---
# Install Packages and Load them
```{r}
#install.packages("haven")
library(haven)
#install.packages("stargazer")
library(stargazer)
#install.packages("labelled")
library(labelled)
#install.packages("plm")
library(plm)
#install.packages("fastDummies") 
library(fastDummies)
```


# Read the Dataset
```{r}
rm(list = ls())

WD_file = dirname(rstudioapi::getSourceEditorContext()$path)
setwd(WD_file)
getwd()
````


```{r}
data_BHI <- haven::read_dta('BDH-Bismarck-JEEA-spec1.dta')
````

Get Variabel Labels
```{r}
var_label(data_BHI)

val_labels(data_BHI)
````

# Explore the Data
```{r}
head(data_BHI)
````

```{r}
summary(data_BHI)
````

```{r}
str(data_BHI$industry)
```

# Check Panel Structure with plm package
Result: Unbalanced Panel: n = 36, T = 336-336, N = 12096
```{r}
pdim(data_BHI)
````

#to be put somewhere / to be used
```{r}
deathrate_year_means <- aggregate(dbo_tot_pc~year, data = data, FUN = mean)

plot(deathrate_year_means)
```

```{r}
library(ggplot2)

# Calculate the mean dbo_tot_pc per year
mean_dbo_tot_pc <- aggregate(dbo_tot_pc ~ year, data = data, FUN = mean)

# Create a scatter plot with the mean dbo_tot_pc per year
ggplot(mean_dbo_tot_pc, aes(x = year, y = dbo_tot_pc)) +
  geom_point() +
  labs(title = "Time Series Plot of Mean dbo_tot_pc",
       x = "Year",
       y = "Mean dbo_tot_pc") +
  theme_minimal()

```

# Figure 1
Long-run development of mortality in Prussia.
The figure shows the crude death rate defined as the total number of deaths per year per 1,000 people.
The vertical bar indicates the introduction of BHI in 1884.
```{r}
library(ggplot2)
# Create a scatter plot of crude death rate over time
ggplot(data_BHI, aes(x = year, y = dbo_tot_pc)) +
  geom_point() +
  geom_vline(xintercept = 1884, linetype = "dashed", color = "red") +
  labs(title = "Long-run Development of Mortality in Prussia",
       x = "Year",
       y = "Crude Death Rate") +
  theme_minimal()

```


# Figure 2
Expansion of Health Insurance in Prussia.
The figure shows the share of health insured in the total population per year.
Data refer to Prussia within its respective borders.
Insurance benefits vary pre- and post-1884.
The vertical bar indicates the introduction of BHI in 1884.



# Figure 3
Crude death rates by sector (blue collar and public servants).
The figure shows district- level mean mortality, in deaths per 1,000,
across blue-collar occupations and public servants.
Following a difference-in-differences logic,
the counterfactual is computed by parallel-shifting the trend
of the public servants up to the blue-collar workers level in 1884.
The vertical bar indicates the introduction of BHI in 1884.
## First about the counterfactual line
```{r}
library(dplyr)

# Generate temporary variables
data_BHI <- data_BHI %>%
  mutate(bluecollar1884 = ifelse(industry == 1 & year == 1884, dbo_tot_pc, NA),
         public1884 = ifelse(industry == 0 & year == 1884, dbo_tot_pc, NA))

# Generate maximum values
dbo_tot_bluecollar1884 <- max(data_BHI$bluecollar1884, na.rm = TRUE)
dbo_tot_public1884 <- max(data_BHI$public1884, na.rm = TRUE)

# Generate counterfactual variable
data_BHI <- data_BHI %>%
  mutate(counterfactual = ifelse(industry == 0 & year >= 1884, 
                                 dbo_tot_pc + dbo_tot_bluecollar1884 - dbo_tot_public1884, NA))

# Drop temporary variables and max value variables
#data_BHI <- data_BHI %>%
  #select(-bluecollar1884, -public1884)

```

```{r}
library(ggplot2)

# Plot the counterfactual variable over time
counterfactual_plot <- ggplot(data_BHI, aes(x = year, y = counterfactual)) +
  geom_line(color = "grey", size = 1.5) +
  labs(title = "Counterfactual Over Time",
       x = "Year",
       y = "Counterfactual") +
  theme_minimal()


counterfactual_plot
````

## Now about the plot

```{r}
library(ggplot2)

# Calculate the mean dbo_tot_pc per year and industry
mean_dbo_tot_pc <- aggregate(dbo_tot_pc ~ year + industry, data = data_BHI, FUN = mean)

# Convert the year variable to numeric
mean_dbo_tot_pc$year <- as.numeric(as.character(mean_dbo_tot_pc$year))

# Create a scatter plot with different colors for each industry level
plot <- ggplot(mean_dbo_tot_pc, aes(x = year, y = dbo_tot_pc, color = factor(industry))) +
  geom_point() +
  geom_line(aes(group = industry), linetype = "dashed") +
  geom_vline(xintercept = 1884, linetype = "dashed", color = "black", size = 1.5) +
  labs(title = "Development of Mean Death Rate by Industry",
       x = "Year",
       y = "Mean dbo_tot_pc") +
  scale_color_manual(values = c("blue", "black"),
                     labels = c("Blue Collar", "Public Servant")) +
  theme_minimal()

plot 
# Save the plot as a PNG file
ggsave("plot.png", plot, width = 8, height = 6, dpi = 300)

```

```{r}
library(ggplot2)

# Calculate the mean dbo_tot_pc per year and industry
mean_dbo_tot_pc <- aggregate(dbo_tot_pc ~ year + industry, data = data_BHI, FUN = mean)

# Convert the year variable to numeric
mean_dbo_tot_pc$year <- as.numeric(as.character(mean_dbo_tot_pc$year))

# Create a scatter plot with different colors for each industry level
plot <- ggplot(mean_dbo_tot_pc, aes(x = year, y = dbo_tot_pc, color = factor(industry))) +
  geom_point() +
  geom_line(aes(group = industry), linetype = "dashed") +
  geom_line(data = data_BHI, aes(x = year, y = counterfactual), color = "grey", size = 1.5) +
  geom_vline(xintercept = 1884, linetype = "dashed", color = "black", size = 1.5) +
  labs(title = "Development of Mean Death Rate by Industry",
       x = "Year",
       y = "Mean dbo_tot_pc") +
  scale_color_manual(values = c("blue", "black"),
                     labels = c("Blue Collar", "Public Servant")) +
  theme_minimal()

# show the plot
plot
````

# First try DiD

## Create dichotomous BlueCollar variable
```{r}
data$BlueCollar <- ifelse(data$industry == 1, 1, 0)
```

Check if BlueCollar variable is correct
```{r}
"BlueCollar" %in% names(data)  # Should return TRUE
is.numeric(data$BlueCollar)  # Should return TRUE
```


```{r}
dbo_tot_pc_did <- plm(dbo_tot_pc ~ industry*factor(year, exclude = 1884) + industry + factor(year) + RB,
                    data = data_BHI,
                    index = c("id_occ", "id_year"), # individual and the time index
                    model = "within", # or "pooling" or "random" or "fd" or "between"
                    effect = "twoways"
                  )

summary(dbo_tot_pc_did)
````

```{r}
# Identify the base coefficient for year 1884
base_coef <- dbo_tot_pc_did[data_BHI$Year == 1884, "Coef"]

# Subtract the base coefficient from all other years' coefficients
base_coef$Normalized_Coef <- dbo_tot_pc_did$Coef - base_coef

# Base R plot with normalized coefficients
plot(data_BHI$Year, base_coef$Normalized_Coef, type='b',
     main='Normalized Year Coefficients', xlab='Year', ylab='Normalized Coefficient')
````

```{r}
# Extract coefficients and their standard errors
model_coef <- coef(summary(dbo_tot_pc_did))

# Create dataframe for the plot
df_coef <- data.frame(
  Year = as.numeric(gsub("industry:factor\\(year\\)", "", rownames(model_coef))),
  Coef = model_coef[, "Estimate"],
  StdErr = model_coef[, "Std. Error"]
)

# Keep only year factors
df_coef <- df_coef[grepl("industry:factor\\(year\\)", rownames(model_coef)), ]

# Normalize coefficients
base_coef <- df_coef[df_coef$Year == 1884, "Coef"]
df_coef$Normalized_Coef <- df_coef$Coef - base_coef

# Calculate confidence intervals (at 95% confidence level)
df_coef$CI_lower <- df_coef$Normalized_Coef - 1.96*df_coef$StdErr
df_coef$CI_upper <- df_coef$Normalized_Coef + 1.96*df_coef$StdErr

# Load the necessary library
library(ggplot2)

# Ggplot with normalized coefficients and confidence intervals
ggplot(df_coef, aes(x=Year, y=Normalized_Coef)) + 
  geom_ribbon(aes(ymin=CI_lower, ymax=CI_upper), alpha=0.2) +  # Confidence intervals
  geom_line() + geom_point() +
  ggtitle('Normalized Year Coefficients with Confidence Intervals') + 
  xlab('Year') + ylab('Normalized Coefficient')
```

```{r}
fixef(dbo_tot_pc_did)
 ````

# test 
```{r}
any(table(data$industry, data$year) > 1)
````


