---
title: "Driva et al. Bismarck Replication"
author: "Till Schickle"
date: "2023-06-02"
output:
  html_document: default
  pdf_document: default
---

# Packages

```{r}
library(tidyverse)
library(haven)
library(plm)
library(modelsummary)
library(lmtest)
library(gridExtra)
library(grid)
library(gridtext)
library(stargazer)
library(webshot)
library(tibble)
library(dplyr)
library(gt)
```


# DiD
## Data

```{r load-libraries & data, message=FALSE, warning=FALSE}

#load data
#pick path
path_data <- file.choose()
df_bismarck_raw <- read_dta(path_data)

#clean data up, rename columns more intuitively
df_bismarck <- df_bismarck_raw %>%
  rename(bluecollar = industry,
         district_id = code,
         district_name = RB,
         deaths_tot = dbo_tot,
         deaths_am = dbo_am,
         deaths_af = dbo_af,
         deaths_km = dbo_km,
         deaths_kf = dbo_kf,
         deaths_tot_pc = dbo_tot_pc,
         deaths_am_pc = dbo_am_pc,
         deaths_af_pc = dbo_af_pc,
         deaths_km_pc = dbo_km_pc,
         deaths_kf_pc = dbo_kf_pc,
         urb_rate = i_urb_pc)

#create panal data set
pdf_bismarck <- pdata.frame(df_bismarck, index = c("id_occ", "id_year"))

#delete unneeded data sets
rm(list = c("df_bismarck_raw", "df_bismarck"))
#delete path var
rm("path_data")
```

## Parallel trends
```{r plotting parallel trends,message=FALSE, warning=FALSE}
#creating df with data for trend plot
trends <- aggregate(deaths_tot_pc~year+bluecollar, data = pdf_bismarck, FUN = mean)
#error message w/o this - unclear why it is necessary
trends$year <- as.numeric(trends$year)
trends$bluecollar <- as.factor(trends$bluecollar)
#assign names instead of numbers
trends$bluecollar <- ifelse(trends$bluecollar==0,"public servant", "blue collar")
#create new rows for counterfactual
trends <- rbind(trends, data.frame(year = seq(1877,1900), 
                                   bluecollar = "blue collar - counterfactual", 
                                   deaths_tot_pc = 0))
#assign the correct values
#up to and including 1884 values are equal to bluecollar
trends$deaths_tot_pc[trends$bluecollar == "blue collar - counterfactual" & 
                      trends$year <= 1884] <- trends$deaths_tot_pc[ 
                        trends$bluecollar == "blue collar" &
                        trends$year <= 1884]

#from 1885 on the values are that of public servant + 
#the difference between bluecollar and public servant in 1884

#calc difference between mean death rate in 1884
diff_1884 = trends$deaths_tot_pc[trends$year == 1884 & trends$bluecollar == "blue collar"] - 
  trends$deaths_tot_pc[trends$year == 1884 & trends$bluecollar == "public servant"]

#assign the values for post 1884
trends$deaths_tot_pc[trends$bluecollar == "blue collar - counterfactual" & 
                      trends$year >= 1885] <- trends$deaths_tot_pc[trends$bluecollar == "public servant" & 
                      trends$year >= 1885] + diff_1884

#plotting
ggplot(trends, aes(x = year, y = deaths_tot_pc, color = bluecollar, group = bluecollar)) +
  geom_line() + 
  #draw blue collar line on top of counterfactual line
  geom_line(data = subset(trends, 
                          bluecollar == "blue collar"), 
                          color = "blue", 
                          size = 1) + 
  labs(x = "Time", y = "Mean Crude Death Rate", title = "Development of Mean Crude Death Rate by Sector") + 
  scale_color_manual(values = c("blue", "gray", "red")) +
  geom_vline(xintercept = 1884, linetype = "dashed", color = "black") + #add the vertical line in 1884
  theme_minimal() + #no gray background
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(linetype = "none") + #deleting legend for linetypes
  theme(legend.position = "bottom",
        legend.title = element_blank()) #adjust design of legend

#delete unneeded df & variable
rm(list = c("diff_1884", "trends"))
```

## Baseline model
```{r,message=FALSE, warning=FALSE}

#to use 1884 as baseline year, setting year to 0 if year is 1884
pdf_bismarck$year_normalized <- ifelse(pdf_bismarck$year == 1884, 
                                      0, 
                                      pdf_bismarck$year)

#run regression
model_did_baseline <- plm(deaths_tot_pc ~ factor(year_normalized) * bluecollar, 
                    data = pdf_bismarck,
                    model = "within")

summary(model_did_baseline)

#extract coefficient estimates
coefficients <- coef(model_did_baseline)
#keep only last 23 entries (relevant estimates of interaction term)
coefficients <- coefficients[-(1:(length(coefficients)-23))]

#clustered standard errors
clustered_sd <- coeftest(model_did_baseline, 
                         vcov=vcovHC(model_did_baseline,
                                     type="HC0",
                                     cluster="group"))
std_errors <- (clustered_sd)[, 2]
#keep only last 23 entries (relevant estimates of interaction term)
std_errors <- std_errors[-(1:(length(std_errors)-23))]

# create data frame with values and names of coefficients
df_coef <- data.frame(coefficient_name = c(1877:1883, 1885:1900), 
                      coefficient = coefficients,
                      standard_error = std_errors)

# insert the row between 1883 and 1885
df_coef <- rbind(df_coef[1:7,], 
                 data.frame(coefficient_name = 1884, 
                           coefficient = 0, 
                           standard_error = 0), 
                 df_coef[8:nrow(df_coef),])


ggplot(data = df_coef,
       aes(x = coefficient_name, 
           y = coefficient)) +
  geom_point() +
  #95% conf. interval
  geom_errorbar(
    aes(ymin = coefficient - 1.96 * standard_error, 
        ymax = coefficient + 1.96 * standard_error), 
    width = 0.2, 
    color = "blue") +
  #90% conf. interval
  geom_errorbar(
    aes(ymin = coefficient - 1.645 * standard_error, 
        ymax = coefficient + 1.645 * standard_error), 
    width = 0.2, 
    color = "grey") +
  theme_minimal() +
  geom_hline(yintercept = 0, 
             linetype = "dashed", 
             color = "red") +
  geom_vline(xintercept = 1884, 
             linetype = "dashed", 
             color = "red") +
  scale_x_continuous(
    breaks = c(1880, 1884, 1890, 1900),
    labels = c("1880", "1884", "1890", "1900")) +
  scale_y_continuous(
    breaks = c(2, 0, -2, -4),
    labels = c(2, 0, -2, -4)) +
  ylim(-4, 2) +
  labs(x = NULL, y = "Coefficient Estimate")

```

## Heterogeneity age & gender
There might be heterogeneous effects for children or women (Figure 5 in paper)

```{r heterogeneity age gender, warning=FALSE, message=FALSE}

'run the baseline model but with different outcome variables: 
death rates of children or women'

variables <- c("deaths_am_pc", "deaths_af_pc", "deaths_km_pc", "deaths_kf_pc")

#loop through the four variables, run regression and create plot for each
for (i in 1:length(variables)) {
  
  #regression formula
  formula <- paste(variables[i], 
                  "factor(year_normalized) * bluecollar", 
                  sep = "~",
                  collapse = "+")
  
  #specify model name
  model_name <- paste("model", 
                      variables[i], 
                      sep = "_", 
                      collapse = "+")
  
  #run the regression
  model <- plm(formula = formula, 
             data = pdf_bismarck,
             model = "within")
  
  #assign name to model
  assign(model_name, model)
  
  #extract coefficient estimates
  coefficients <- coef(model)
  #keep only last 23 entries (relevant estimates of interaction term)
  coefficients <- coefficients[-(1:(length(coefficients)-23))]
  
  #clustered standard errors
  clustered_sd <- coeftest(model, 
                           vcov=vcovHC(model,
                                       type="HC0",
                                       cluster="group"))
  std_errors <- (clustered_sd)[, 2]
  #keep only last 23 entries (relevant estimates of interaction term)
  std_errors <- std_errors[-(1:(length(std_errors)-23))]
  
  # create data frame with values and names of coefficients
  df_coef <- data.frame(coefficient_name = c(1877:1883, 1885:1900), 
                        coefficient = coefficients,
                        standard_error = std_errors)
  
  # insert the row between 1883 and 1885
  df_coef <- rbind(df_coef[1:7,], 
                   data.frame(coefficient_name = 1884, 
                             coefficient = 0, 
                             standard_error = 0), 
                   df_coef[8:nrow(df_coef),])

  #create plot
  plot <- ggplot(data = df_coef,
                 aes(x = coefficient_name, 
                     y = coefficient)) + 
    geom_point() +
    #95% conf. interval
    geom_errorbar(
      aes(ymin = coefficient - 1.96 * standard_error, 
          ymax = coefficient + 1.96 * standard_error), 
      width = 0.2, 
      color = "blue") +
    theme_minimal() +
    geom_hline(yintercept = 0, 
               linetype = "dashed", 
               color = "red") +
    geom_vline(xintercept = 1884, 
               linetype = "dashed", 
               color = "red") +
    scale_x_continuous(
      breaks = c(1880, 1884, 1890, 1900),
      labels = c("1880", "1884", "1890", "1900")) +
    scale_y_continuous(
      breaks = c(2, 0, -2, -4),
      labels = c(2, 0, -2, -4)) +
    ylim(-4, 2) +
    theme(plot.title = element_text(hjust = 0.5)) + 
    labs(x = NULL, y = NULL)
  
  #name plot
  plot_name <- paste("plot", 
                     variables[i],
                     sep = "_", 
                     collapse = "+")
  
  #assign name to plot
  assign(plot_name, plot)
  
}
plot_deaths_am_pc <- plot_deaths_am_pc + labs(title = "Adult Males")
plot_deaths_af_pc <- plot_deaths_af_pc + labs(title = "Adult Females")
plot_deaths_km_pc <- plot_deaths_km_pc + labs(title = "Boys < 14")
plot_deaths_kf_pc <- plot_deaths_kf_pc + labs(title = "Girls < 14")

#create four plots in a grid to show heterogeneous effects
combined_plot <- grid.arrange(plot_deaths_am_pc, 
                              plot_deaths_af_pc, 
                              plot_deaths_km_pc,
                              plot_deaths_kf_pc, 
                              nrow = 2, 
                              ncol = 2,
                              left = richtext_grob("Coefficient Estimates", rot = 90))

ggsave("heterogeneity_gender_age.png", combined_plot)


```

## Heterogeneity industries

The effects might be confounded by wage effects in the industries. A consistent effect across all industries would serve as evidence against this caveat.


```{r}

  
  pdf_bismarck$occ_normalized <- ifelse(pdf_bismarck$occ == "14", 
                                      0, 
                                      pdf_bismarck$occ)
  
  #run regression
  model_did_heterogeneity_industries <- plm(deaths_tot_pc ~ factor(year_normalized) * factor(occ_normalized), 
                      data = pdf_bismarck,
                      model = "within")

  #extract coefficient estimates
  coefficients <- coef(model_did_heterogeneity_industries)
  
  #clustered standard errors
  clustered_sd <- coeftest(model_did_heterogeneity_industries, 
                           vcov=vcovHC(model_did_heterogeneity_industries,
                                       type="HC0",
                                       cluster="group"))
  
  #names of industries
  industries = c("Mining & turf", "Minerals", "Metals", "Machinery", 
                 "Chemicals", "Fossil fuels", "Textiles", "Paper & leather", 
                 "Wood", "Food", "Appareal & cleaning", "Construction", 
                 "Printing")
  
  
  for (i in 1:13) {
    #calculate where the relevant coefficients and std. errors are in the vectors
    start = i*23+1
    end = i*23+23
    #crate data frame with coefficients
    df <- data.frame(coefficient_name = c(1877:1883, 1885:1900),
                     coefficient = coefficients[start:end],
                     standard_error = clustered_sd[start:end,2])
    
    # insert the row between 1883 and 1885
    df <- rbind(df[1:7,], 
                data.frame(coefficient_name = 1884, 
                           coefficient = 0, 
                           standard_error = 0), 
                     df[8:nrow(df),])
    
    #create plot
    plot <- ggplot(data = df,
                 aes(x = coefficient_name, 
                     y = coefficient)) + 
      geom_point(size=0.5) +
      #95% conf. interval
      geom_errorbar(
        aes(ymin = coefficient - 1.96 * standard_error, 
            ymax = coefficient + 1.96 * standard_error), 
        width = 0, 
        color = "blue") +
      theme_minimal() +
      theme(plot.title = element_text(size = 10, hjust = 0.5)) +
      geom_hline(yintercept = 0, 
                 linetype = "dashed", 
                 color = "red") +
      geom_vline(xintercept = 1884, 
                 linetype = "dashed", 
                 color = "red") +
      ylim(-11, 7) +
      labs(x=NULL, y=NULL, title = industries[i]) +
      scale_x_continuous(
        breaks = c(1879, 1884, 1889, 1894, 1899),
        labels = c("79", "84", "89", "94", "99"))
    
    #assign name to plot
    plot_name = paste("plot",
                      industries[i],
                      sep = "_",
                      collapse = "+")
    assign(plot_name, plot)
    
  }
  
  #create four plots in a grid to show heterogeneous effects
  combined_plot <- grid.arrange(`plot_Mining & turf`,
               plot_Minerals,
               plot_Metals,
               plot_Machinery,
               plot_Textiles,
               `plot_Paper & leather`,
               plot_Wood,
               plot_Food,
               `plot_Appareal & cleaning`,
               plot_Construction,
               ncol = 5,
               nrow = 2,
               left = richtext_grob("Coefficient Estimates", rot = 90))
  
  
  ggsave("heterogeneity_industry.png", combined_plot, width = 15, height = 8, units = "cm")
  

```


# County Fixed Effects

Main issue with DiD is that selection into occupation groups after 1884 cannot be excluded.
Both selection that would lead to upward bias (sicker, more care needing people select into bluecollar occupations to get insurance) as well as selection leading to downward bias (younger people select into bluecollar occupations as the job becomes more attractive due to insurance) are possible.
Thus, need different approach such that one can fix treatment assignment/intensity pre treatment: county fixed effects.

## Data

```{r}
#load data
#pick path for spec2 data
path_data <- file.choose()
df_bismarck_raw <- read_dta(path_data)

#clean data up, rename columns more intuitively
df_bismarck <- df_bismarck_raw %>%
  rename(period1875 = yr1,
         period1880 = yr2,
         period1885 = yr3,
         period1890 = yr4,
         period1895 = yr5,
         period1900 = yr6,
         treat1875 = b_industry_w_pc1875,
         treat1880 = b_industry_w_pc1880,
         treat1885 = b_industry_w_pc1885,
         treat1890 = b_industry_w_pc1890,
         treat1895 = b_industry_w_pc1895,
         treat1900 = b_industry_w_pc1900,
         deaths_tot = dth_pc,
         county = code1867,
         bluecollar_1882 = b_industry_w_pc,
         deaths_male = dthm_pc,
         deaths_female = dthf_pc,
         deaths_infants_tot = dth1_pb,
         deaths_infants_leg = dthleg_pb,
         deaths_infants_illeg = dthbas_pb,
         bluecollar_1882_self = b_industry_self_pc,
         public_1882 = e_public_w_pc)

#create panal data set
pdf_bismarck <- pdata.frame(df_bismarck, index = c("county", "year"))

#delete unneeded data sets
rm(list = c("df_bismarck_raw", "df_bismarck"))
#delete path var
rm("path_data")


```


## Model

```{r}

  # create new year variable which is 0 whenever year = 1880 as this is the 
  # omitted reference period
  pdf_bismarck$year_normalized <- ifelse(pdf_bismarck$year == 1880, 
                                         0, 
                                         pdf_bismarck$year)
  

  ### run regressions
  # baseline fixed effects model
  model_fe_baseline <- plm(formula = deaths_tot ~ 
                             bluecollar_1882*factor(year_normalized),
                           data = pdf_bismarck,
                           model = "within")
  
  # create list of models and add baseline model
  models <- list("Baseline" = model_fe_baseline)

  
  # list of dep. variables for models w/ controls
  dep_vars <- c("deaths_tot",
                "deaths_male",
                "deaths_female",
                "deaths_infants_tot",
                "deaths_infants_leg",
                "deaths_infants_illeg")
  
  # list of names for models
  model_names <- c("Controls",
                   "Males",
                   "Females",
                   "Infants",
                   "Legitimate Infants",
                   "Illegitimate Infants")
  
  # loop that runs the models w/ controls
  for (i in 1:6) {
    
    formula <- paste(dep_vars[i], 
                     "bluecollar_1882*factor(year_normalized) " %>%
                       paste0("+ urb_pc + waterwork_pc + sewage_pc"),
                     sep = "~")
    
    model <- plm(formula = formula,
                 data = pdf_bismarck,
                 model = "within")
    
    
    models[[length(models) + 1]] <- model
    names(models)[length(models)] <- model_names[i]
    
  }
  
  # placebo tests
  # self-employed
  model_fe_placebo_self <- plm(formula = deaths_tot ~ 
                             bluecollar_1882_self*factor(year_normalized)+ 
                               urb_pc + waterwork_pc + sewage_pc,
                           data = pdf_bismarck,
                           model = "within")
  summary(model_fe_placebo_self)
  
  # create list of models and add baseline model
  models[[length(models) + 1]] <- model_fe_placebo_self
  names(models)[length(models)] <- "Placebo Self-Employed"

  # public servants
  model_fe_placebo_public <- plm(formula = deaths_tot ~ 
                             public_1882*factor(year_normalized)+ 
                               urb_pc + waterwork_pc + sewage_pc,
                           data = pdf_bismarck,
                           model = "within")
  
  # create list of models and add baseline model
  models[[length(models) + 1]] <- model_fe_placebo_public
  names(models)[length(models)] <- "Placebo Public Servants"
  
  # clustered standard erros
  # function that calculates the clustered standard errors for the passed model 
  vcov_sss_group <- function(model) {
    vcovHC(model, type = "sss", cluster = "group")
  }
  
    
  # new 
  rows <- tribble(~term, ~`[[1]]`, ~`[[2]]`, ~`[[3]]`, ~`[[4]]`, ~`[[5]]`, ~`[[6]]`, ~`[[7]]`, ~`[[8]]`, ~`[[9]]`,
                  'Controls', 'No', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes',
                  'County FE', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes',
                  'Time FE', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes', 'Yes',
                  'Counties', '441', '441', '441', '441', '441', '441', '441', '441', '441',
                  'Periods', '6', '6', '6', '6', '6', '6', '6', '6', '6')

  attr(rows, 'position') <- c(11, 12, 13, 15, 16)
  
  
  # create regression table 
  reg_table <- modelsummary(models,
               vcov = vcov_sss_group,
               #name all three different sets of coefficients the same such 
               #that the table consists of only five rows of coefficients
               coef_map = c('bluecollar_1882:factor(year_normalized)1' = 'Treat*1875', 
                            'bluecollar_1882:factor(year_normalized)3' = 'Treat*1885',
                            'bluecollar_1882:factor(year_normalized)4' = 'Treat*1890',
                            'bluecollar_1882:factor(year_normalized)5' = 'Treat*1895',
                            'bluecollar_1882:factor(year_normalized)6' = 'Treat*1900',
                            'bluecollar_1882_self:factor(year_normalized)1' = 'Treat*1875', 
                            'bluecollar_1882_self:factor(year_normalized)3' = 'Treat*1885',
                            'bluecollar_1882_self:factor(year_normalized)4' = 'Treat*1890',
                            'bluecollar_1882_self:factor(year_normalized)5' = 'Treat*1895',
                            'bluecollar_1882_self:factor(year_normalized)6' = 'Treat*1900',
                            'public_1882:factor(year_normalized)1' = 'Treat*1875', 
                            'public_1882:factor(year_normalized)3' = 'Treat*1885',
                            'public_1882:factor(year_normalized)4' = 'Treat*1890',
                            'public_1882:factor(year_normalized)5' = 'Treat*1895',
                            'public_1882:factor(year_normalized)6' = 'Treat*1900'),
               gof_map = c("nobs", "r.squared"),
               add_rows = rows,
               output = "gt")
  
  reg_table <- reg_table %>%

    # column labels
    tab_spanner(label = 'Total Deathrate', columns = 2:3) %>%
    tab_spanner(label = 'Heterogeneity Tests', columns = 4:8) %>%
    tab_spanner(label = 'Placebo Tests', columns = 9:10)
  
  reg_table
  
  gtsave(reg_table, file = "fe_reg_table.docx")

```

## Robustness checks

```{r}

```

